An essentially algebraic formulation of dependent type theory

### 2015-07-29 main.tex
separation of sections into different files

\begin{abstract}
We present an algebraic formulation of dependent type theory with
three sorts for contexts, families and terms, with finitely many operations
between them and finitely many equality axioms. Algebras for this theory are
called E-systems. We then show that there is a full and faithful functor
from Voevodsky's B-systems into E-systems, of which the image is the category
of stratified E-systems. 
\end{abstract}

@@

@@

\begin{rmk}
We have chosen not to use common type theoretical notation involving special
expressions for judgments and for inference rules. One reason is that it has been
proven to be hard for the uninitiated to read and comprehend such expressions.
Another is that set theoretic
notation serves very well for getting the algebraic structure accross, which
is what we want to focus on.

@@

\input{weakeningcats}

>>>>

\begin{abstract}
We present an algebraic formulation of dependent type theory with
three sorts for contexts, families and terms, with finitely many operations
between them and finitely many equality axioms. Algebras for this theory are
called E-systems. A distinctive feature of the dependent type theory of
E-systems is that it takes `telescopic families' as the primitive notion of
families.

We then compare Voevodsky's B-systems, Dybjer's categories
with families and Jacob's comprehension categories with E-systems. There is
a full and faithful functor from B-systems to E-systems, of which the image is
the category of stratified E-systems. To have a functor from categories with
families to E-systems, we need to impose an additonal coherence condition, which
makes sure that the substitution operation of a category with families extends 
to telescopic families. 
\end{abstract}

@@

\subsection*{Principles of mathematical fuseki}
Mathematics is the study of abstract concepts such as space or quantity, and
when we design a language for such a study, we want to be able to efficiently
express things about the objects under investigation. The subject matter of
this paper is a logical framework of dependent type theory.

Pfenning \cite{PfenningLFintro} defines logical frameworks
to be meta-languages for the formalization of deductive system. Examples of
logical framework therefore include propositional logic, first-order predicate 
logic, Russell's ramified type theory, the lambda-calculus, dependent type 
theory, and so on. By considering models of a logical framework, it becomes
itself the subject of mathematical study.

It may actually be useful to see logical frameworks themselves
as objects subject to mathematical investigation.

Logical frameworks for mathematics are not set in stone.

 and neither to they
remain the same once they are declared.

When designing and maintaining a logical framework

@@

\begin{rmk}
We have chosen not to use common type theoretical notation involving special
expressions for judgments and for inference rules, because we want to focus
on the algebraic structure of type dependency rather than on its syntax.

@@

\input{e-systems}

\input{internal_homs}

\input{compared}

\section{Type constructors in E-systems}

\begin{defn}
An E-system $\mathbb{E}$ is said to have $\Pi$-types, if 
\begin{enumerate}
\item for any 
$A\in\cat{F}/\Gamma$, there is an E-homomorphism
\begin{equation*}
\Pi_A : \mathbb{E}/\ctxext{\Gamma}{A}\to\mathbb{E}/\Gamma
\end{equation*}
For terms $g:Q$ in context $\ctxext{{\Gamma}{A}}{P}$, we will usually write
$\lambda_{A}(g)$ for the term $\Pi_A(g)$ of $\Pi_A(Q)$ in context
$\ctxext{\Gamma}{\Pi_A(P)}$. Note that when $P$ is empty, then we retrieve the
usual lambda-abstraction.
\item for any $A\in\cat{F}/\Gamma$ and $Q\in\cat{F}/\ctxext{{\Gamma}{A}}{P}$, 
there is a term 
\begin{equation*}
\tfev^{A,Q} : \ctxwk{\Pi_A(Q)}{{\Pi_A(P)}{Q}}
\end{equation*}
in context $\ctxext{{{{\Gamma}{\Pi_A(P)}}{\Pi_A(Q)}}{\ctxwk{\Pi_A(\ctxext{P}{Q})}{A}}}{\ctxwk{\Pi_A(\ctxext{P}{Q})}{P}}$.
Note that for each $g:\Pi_A(Q)$, we obtain a term $\subst{g}{\tfev^{A,Q}}: \ctxwk{\Pi_A(P)}{Q}$ in
context $\ctxext{{\Gamma}{\Pi_A(P)}}{\ctxwk{\Pi_A(P)}{A}}$.
\item for any $A\in\cat{F}/\Gamma$ there is a term
\begin{equation*}
{\tfev^{A,P}} : {\ctxwk{\Pi_A(P)}{P}}
\end{equation*}
in context ${\ctxext{{\Gamma}{\Pi_A(P)}}{\ctxwk{\Pi_A(P)}{A}}}$
\end{enumerate}
satisfying
\begin{enumerate}
\item 
\begin{enumerate}
\item For any $P\in\cat{F}/\ctxext{\Gamma}{A}$ one has
\begin{equation*}
\Pi_{\ctxext{A}{P}} \jdeq \Pi_A\circ \Pi_P
\end{equation*}
\item For any $Q\in\cat{F}/\ctxext{{\Gamma}{A}}{P}$, one has
\begin{equation*}
{\evtm{\ctxext{A}{P}}{Q}}
  \jdeq
{\subst{\evtm{A}{\mprd{P}{Q}}}{\ctxwk{\mprd{A}{\mprd{P}{Q}}}{\evtm{P}{Q}}}}
\end{equation*}
\end{enumerate}
\item 
\begin{enumerate}
\item For any $\Gamma\in\mathbb{E}$, one has
\begin{equation*}
\Pi_{\catid{\Gamma}}\jdeq \catid{\mathbb{E}/\Gamma}.
\end{equation*}
\item For any $A\in\mathbb{F}/\Gamma$, one has
\begin{equation*}
\tfev^{\catid{\Gamma},A} \jdeq \tfid{A}
\end{equation*}
\end{enumerate}
\item for any term $f:P$ in context $\ctxext{\Gamma}{A}$, one has
\begin{equation*}
\subst{\lambda_A(f)}{\tfev^{A,P}}\jdeq f.
\end{equation*}
\end{enumerate}
\end{defn}

\begin{rmk}
The evaluation term determines a term
\begin{equation*}
\tilde\tfev^{A,P} : \thom{\ctxwk{A}{\Pi_A(P)}}{P}
\end{equation*}
\end{rmk}

\begin{defn}
A $\Pi$E-homomorphism from $\mathbb{E}$ to $\mathbb{E}'$ is an E-homomorphism
$H:\mathbb{E}\to\mathbb{E}'$ such that
\begin{enumerate}
\item for any $A\in\cat{F}/\Gamma$, one has
\begin{equation*}
\begin{tikzcd}[column sep=large]
\cat{F}/\ctxext{\Gamma}{A} \arrow[d,swap,"\Pi_A"] \arrow[r,"H/\ctxext{\Gamma}{A}"] & \cat{F}'/\ctxext{H(\Gamma)}{H(A)} \arrow[d,"\Pi_{H(A)}"] \\
\cat{F}/\Gamma \arrow[r,swap,"H/\Gamma"] & \cat{F}'/F(\Gamma)
\end{tikzcd}
\end{equation*}
\item For any $P\in\cat{F}/\ctxext{\Gamma}{A}$, one has
\begin{equation*}
H(\tfev^{A,P})\jdeq \tfev^{H(A),H(P)}
\end{equation*}
\end{enumerate}
\end{defn}

### 2015-07-29 weakeningcats.tex
separation of sections into different files

In this section we recall Dybjer's notion of categories with families and we
introduce the equivalent notion of D-systems. After having proved that categories with families
are the same things as D-systems, we will focus on constructing a functor from
D-systems to E-systems without empty context. However, due to the existence of pathological examples
of categories with families, this functor is not going to be faithful. We then
introduce the notion of well-founded D-systems, and show that the
functor from D-systems to E'-systems without empty context restricts to an
embedding from the well-founded D-systems to E'-systems.

@@

\begin{constr}
We construct the asserted structure by
induction on the length of the arrow $A$ into $\Gamma$. We define
$\Lambda(f,\emptyf[\Gamma])\defeq \emptyf[\Delta]$, and $\pi_2(f,\emptyf[\Gamma])\defeq f:\ctxext{\Delta}{\catid{\Delta}}\to\ctxext{\Gamma}{\catid{\Gamma}}$. The resulting square is trivially
a pullback square. Now suppose we have a pullback square as above.
and let $B\in\mathrm{Ty}(\ctxext{\Gamma}{A})$. Then we define
\begin{equation*}
\Lambda(f,[A,B])\defeq [\Lambda(f,A),\pi_2(f,A)_\ast(B)].
\end{equation*}
By \autoref{thm:DD}, the top square in the diagram
\begin{equation*}
\begin{tikzcd}[column sep=6em]
\ctxext{{\Delta}{\Lambda(f,A)}}{(\pi_2(f,A)_\ast(B))} \arrow[d,fib] \arrow[r,"{\pi_2(\pi_2(f,A),B)}"] & \ctxext{{\Gamma}{A}}{B} \arrow[d,fib] \\
\ctxext{\Delta}{\Lambda(f,A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
\Delta \arrow[r,"f"] & \Gamma
\end{tikzcd}
\end{equation*}
is a pullback square.
Since the bottom square is assumed to be a pullback, we get that the outer rectangle
is a pullback, as desired. 
\end{constr}

\begin{rmk}
The construction above does not help us define a functor $f_\ast:\cat{F}/\Gamma
\to\cat{F}/\Delta$ without further assumptions on the D-system $\mathbb{D}$.
The functor $f_\ast$ would have the property that the square
\begin{equation*}
\begin{tikzcd}
\Lambda(\Gamma) \arrow[d,swap,"p"] \arrow[r,"\Lambda(f)"] & \Lambda(\Delta) \arrow[d,"p"] \\
\cat{F}/\Gamma \arrow[r,densely dotted,"{f_\ast}"] & \cat{F}/\Delta
\end{tikzcd}
\end{equation*}
commutes. The above square can only commute, if one has
\begin{equation*}
\Lambda(f)/A\jdeq\Lambda(f)/A'
\end{equation*} 
for any $A,A'\in\Lambda(\Gamma)$ with $p_A\jdeq p_{A'}$. This would entail that
$\pi_2(f,A)\jdeq\pi_2(f,A')$, a property that seems to not hold generally.
\end{rmk}

\begin{defn}
A D-system is said to be coherent if for any $f:\Delta\to\Gamma$ and any
$A,A'\in\Lambda(\Gamma)$ satisfying $p_A\jdeq p_{A'}$, one has
\begin{equation*}
\pi_2(f,A)\jdeq \pi_2(f,A').
\end{equation*}
\end{defn}

@@

\begin{defn}
For any locally well-founded D-system $\mathbb{D}$ there is 
an E'-system $U(\mathbb{D})$ without terminal object in 
$\cat{F}$. Thus, we get a functor $\mathbf{Dsys}\to\mathbf{E'sys}_{\circ}$
\end{defn}

\begin{constr}
Let $f:\Delta\to\Gamma$. By the assumption that $\mathbb{D}$ is locally well-founded,

To verify that this gives indeed a functor from $\cat{F}/\Gamma$ to
$\cat{F}/\Delta$, we need to show that for any two objects or $A,A'\in\Lambda(\Gamma)$
which get mapped to the same object of $\cat{F}/\Gamma$, get mapped to the
same object of $\cat{F}/\Delta$. Let $A,A'\in\Lambda(\Gamma)$ be such that
$p_A\jdeq p_{A'}$. It suffices to show that $\pi_2(f,A)\jdeq\pi_2(f,A')$

The inductive proofs to verify properties
\autoref{defn:E'sys:b} and \autoref{defn:E'sys:c} of \autoref{defn:E'sys} are
completely straightforward, so $U(\mathbb{D})$ is an E'-system without terminal
object in $\cat{F}$.
\begin{comment}
Finally, we need to verify that the structure we've defined satisfies the properties
of being an E'-system without terminal object. We go through them systematically,
following the same order as in \autoref{defn:E'sys}:
\begin{enumerate}[label=(\alph*)]
\item This property holds by definition.
\item Let $\Gamma\in\cat{C}$. We will prove by induction on $A\in\cat{F}/\Gamma$
that $(\catid{\Gamma})_\ast(A)\jdeq A$ and that $\pi_2(\catid{\Gamma},A)\jdeq \catid{\ctxext{\Gamma}{A}}$.
Note that these properties hold by definition for $A\jdeq\catid{\Gamma}$. For the
inductive step, assume that $A\in\cat{F}/\Gamma$ is such that
$(\catid{\Gamma})_\ast(A)\jdeq A$ and $\pi_2(\catid{\Gamma},A)\jdeq
\catid{\ctxext{\Gamma}{A}}$ and let $B\in\mathrm{Ty}(\ctxext{\Gamma}{A})$. Then
we have
\begin{align*}
(\catid{\Gamma})_\ast(A\circ p_B) & \jdeq (\catid{\Gamma})_\ast(A)\circ \pi_2(\catid{\Gamma},A)_\ast(p_B) \\
& \jdeq A\circ p_B
\end{align*}
Also, we have
\begin{align*}
\pi_2(\catid{\Gamma},A\circ p_B) & \jdeq \pi_2(\pi_2(\catid{\Gamma},A),p_B) \\
& \jdeq \pi_2(\catid{\ctxext{\Gamma}{A}},p_B) \\
& \jdeq \catid{\ctxext{{\Gamma}{A}}{B}}
\end{align*}
\item Let $f:\Delta\to \Gamma$ and $g:X\to \Delta$. We will prove by induction
on $A\in\cat{F}/\Gamma$ that $(f\circ g)_\ast(A)\jdeq g_\ast(f_\ast(A))$ and
that $\pi_2(f\circ g,A)\jdeq \pi_2(f,A)\circ \pi_2(g,f_\ast(A))$. Note that
these properties hold by definition for $A\jdeq\catid{\Gamma}$. For the inductive
step, assume that $A\in\cat{F}/\Gamma$ is such that $(f\circ g)_\ast(A)\jdeq g_\ast(f_\ast(A))$ and
$\pi_2(f\circ g,A)\jdeq \pi_2(f,A)\circ \pi_2(g,f_\ast(A))$, and let
$B\in\mathrm{Ty}(\ctxext{\Gamma}{A})$. Then we have
\begin{align*}
(f\circ g)_\ast(A\circ p_B) & \jdeq \pi_2(f\circ g,A)_\ast(p_B) \\
& \jdeq (\pi_2(f,A)\circ\pi_2(g,f_\ast(A)))_\ast(p_B) \\
& \jdeq \pi_2(g,f_\ast)_\ast(\pi_2(f,A)_\ast(p_B)) \\
& \jdeq g_\ast(f_\ast(A\circ p_B))
\end{align*}

\item This property holds by definition.
\end{enumerate}
\end{comment}
\end{constr}

\begin{thm}
The functor $U:\mathbf{Dsys}\to\mathbf{E'sys_{\circ}}$ is full and faithful.
\end{thm}

\begin{proof}
The functor $U$ is faithful, because for any morphism 
$\mathcal{F}:\mathbb{D}\to\mathbb{D}'$ of locally well-founded D-systems, 
$U(\mathcal{F})$ and $\mathcal{F}$ have the same underlying functor
from $\cat{C}$ to $\cat{C}'$.

It remains to show that $U$ is full. Let
$\mathbb{D}$ and $\mathbb{D}$ be D-systems and let $\mathcal{G}:U(\mathbb{D})
\to U(\mathbb{D}')$. Then $\mathcal{G}$ restricted to $\cat{F}$ lifts factorizations.
We  first have to define a natural transformation
$\mathrm{Ty}\Rightarrow \mathrm{Ty}'\circ\mathcal{G}$. Let
$\Gamma\in\cat{C}$ and $A\in\mathrm{Ty}(\Gamma)$. Then we have
$\mathcal{G}(p_A):\mathcal{G}(\ctxext{\Gamma}{A})\to \mathcal{G}(\Gamma)$.
The category of factorizations of $p_A$ in $\cat{F}$ is a isomorphic to $\mathbf{2}$,
so the category of factorizations of $\mathcal{G}(p_A)$ is also isomorphic to
$\mathbf{2}$. This implies that $\mathcal{G}(p_A)$ is of the form $p_B$ for
some $B\in\mathrm{Ty}'(\mathcal{G}(\Gamma))$. This $B$ is called $\mathcal{G}(A)$.

>>>>

In this section we recall Dybjer's notion of categories with families and we
introduce the equivalent notion of D-systems. After having proved that categories with families
are the same things as D-systems, we will focus on constructing a canonical functor from
D-systems to E'-systems, in which the category $\cat{F}$ of families has contexts
as objects and compositions of projections as morphisms. There is no reason why
this category should possess a terminal object, so we naturally have to retreat
to E'-systems without empty contex. However, the construction of a
canonical functor in which families are represented by compositions of projections
is still not possible without a further coherence condition. To make sure that
the pullback operations of a D-system extend to the category
$\cat{F}$, we introduce \emph{coherent} D-systems. So we only have a functor $U$
from the category $\mathbf{Dsys}$ of coherent D-systems to the category
$\mathbf{E'sys_{\circ}}$ of E'-systems without empty context.
We see that it is not evident at all to turn categories with families into
E-systems, even if we ignore the issue that categories with families map
naturally into E'-systems without empty context.
Thus it seems that the type theories of categories with families and of E-systems
are not directly comparable.

We end this section by showing that the functor 
$U:\mathbf{Dsys}\to\mathbf{E'sys_{\circ}}$ restricts to a full and faithful 
functor on the coherent D-systems which satisfy the condition that every 
composition of projections is a composition of projections in precisely one way.
We call coherent D-systems with this property
locally well-founded. Among the locally well-founded D-systems are the B-systems.
As a by-product, we get that the full and faithful functor from B-systems to
E-systems factors through D-systems.

@@

\begin{constr}
We construct the asserted structure by
induction on the length of the arrow $A$ into $\Gamma$. We define
$\Lambda(f)$ on the empty sequence as
$\Lambda(f,\emptyf[\Gamma])\defeq \emptyf[\Delta]$, and we define
$\pi_2(f,\emptyf[\Gamma])\defeq f:\ctxext{\Delta}{\catid{\Delta}}\to
\ctxext{\Gamma}{\catid{\Gamma}}$. The resulting square is trivially
a pullback square. Now suppose we have a pullback square as above.
and let $B\in\mathrm{Ty}(\ctxext{\Gamma}{A})$. Then we define
\begin{equation*}
\Lambda(f,[A,B])\defeq [\Lambda(f,A),\pi_2(f,A)_\ast(B)].
\end{equation*}
By \autoref{thm:DD}, the top square in the diagram
\begin{equation*}
\begin{tikzcd}[column sep=6em]
\ctxext{{\Delta}{\Lambda(f,A)}}{(\pi_2(f,A)_\ast(B))} \arrow[d,fib] \arrow[r,"{\pi_2(\pi_2(f,A),B)}"] & \ctxext{{\Gamma}{A}}{B} \arrow[d,fib] \\
\ctxext{\Delta}{\Lambda(f,A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
\Delta \arrow[r,"f"] & \Gamma
\end{tikzcd}
\end{equation*}
is a pullback square.
Since the bottom square is assumed to be a pullback, we get that the outer rectangle
is a pullback, as desired. 
\end{constr}

\begin{rmk}
We can't directly use the functors $\Lambda(f)$ to give a D-system the structure
of an E'-system, because there is a coherence condition missing.
\end{rmk}

\begin{thm}\label{thm:Dsys_coh}
The following are equivalent:
\begin{enumerate}
\item The functors $\Lambda(f):\Lambda(\Gamma)\to\Lambda(\Delta)$ extend to 
the structure of an E'-system without empty context.
\item For any $f:\Delta\to\Gamma$ and all $A,A'\in\Lambda(\Gamma)$ satisfying 
$p_A\jdeq p_{A'}$, one has
\begin{equation*}
p_{\Lambda(f,A)}\jdeq p_{\Lambda(f,A')}\qquad\text{and}\qquad\pi_2(f,A)\jdeq \pi_2(f,A').
\end{equation*}
\end{enumerate}
\end{thm}

\begin{proof}
Suppose we can extend $\Lambda(f)$ for all $f:\Delta\to\Gamma$, to form commuting
diagrams
\begin{equation*}
\begin{tikzcd}
\Lambda(\Gamma) \arrow[d,swap,"p"] \arrow[r,"\Lambda(f)"] & \Lambda(\Delta) \arrow[d,"p"] \\
\cat{F}/\Gamma \arrow[r,densely dotted,"{f_\ast}"] & \cat{F}/\Delta
\end{tikzcd}
\end{equation*}
of functors, forming an E'-system without empty context. Let 
$f:\Delta\to\Gamma$, and let $A,A':\Lambda(\Gamma)$ be such that 
$p_A\jdeq p_{A'}$. Then it follows
that $\pi_2(f,A)_\ast\jdeq f_\ast/A\jdeq f_\ast/A'\jdeq\pi_2(f,A')_\ast$, so
we get the desired equality $\pi_2(f,A)\jdeq\pi_2(f,A')$.

For the converse, suppose that for all $f:\Delta\to\Gamma$ and $A,A'\in
\Lambda(\Gamma)$ such that $p_A\jdeq p_{A'}$, one has 
$p_{\Lambda(f,A)}\jdeq p_{\Lambda(f,A')}$ and $\pi_2(f,A)\jdeq\pi_2(f,A')$.
Let $f:\Delta\to\Gamma$. By the assumption that $p_{\Lambda(f,A)}\jdeq p_{\Lambda(f,A')}$,
it follows that the objects part of $p^\Delta\circ\Lambda(f):\Lambda(\Gamma)\to\cat{F}/\Delta$ 
factors through $p^\Gamma:\Lambda(\Gamma)\to\cat{F}/\Gamma$. Moreover, by the
assumption that $\pi_2(f,A)\jdeq \pi_2(f,A')$ it follows that the pullback
squares
\begin{equation*}
\begin{tikzcd}[column sep=large]
\ctxext{\Delta}{f_\ast(A)} \arrow[d,fib] \arrow[r,"{\pi_2(f,A)}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
\Delta \arrow[r,"f"] & \Gamma
\end{tikzcd}
\qquad\text{and}\qquad
\begin{tikzcd}[column sep=large]
\ctxext{\Delta}{f_\ast(A')} \arrow[d,fib] \arrow[r,"{\pi_2(f,A')}"] & \ctxext{\Gamma}{A} \arrow[d,fib] \\
\Delta \arrow[r,"f"] & \Gamma
\end{tikzcd}
\end{equation*}
are the same. Since we have this property
for any object $\Gamma$, it also follows that the morphisms part of
$p^\Delta\circ\Lambda(f)$ factors through $p^\Gamma$. This defines the functor
$f_\ast$ in the commuting square
\begin{equation*}
\begin{tikzcd}
\Lambda(\Gamma) \arrow[d,swap,"p"] \arrow[r,"\Lambda(f)"] & \Lambda(\Delta) \arrow[d,"p"] \\
\cat{F}/\Gamma \arrow[r,densely dotted,"{f_\ast}"] & \cat{F}/\Delta
\end{tikzcd}
\end{equation*}
We need to verify that $f_\ast$ is a functorial choice of pullbacks, as in the
definition of \autoref{defn:E'sys}. Since $f_\ast$ is defined as an extension
of $\Lambda(f)$, it is enough to verify the analoguous functoriality properties
for $\Lambda(f)$. We do this by induction on the objects of $\Lambda(\Gamma)$,
in the same order as in \autoref{defn:E'sys}.
\begin{enumerate}[label=(\alph*)]
\item This property holds by definition.
\item Let $\Gamma\in\cat{C}$. We will prove by induction on $A\in\Lambda(\Gamma)$
that $(\catid{\Gamma})_\ast(A)\jdeq A$ and that $\pi_2(\catid{\Gamma},A)\jdeq \catid{\ctxext{\Gamma}{A}}$.
Note that these properties hold by definition for $A\jdeq\catid{\Gamma}$. For the
inductive step, assume that $A\in\Lambda(\Gamma)$ is such that
$(\catid{\Gamma})_\ast(A)\jdeq A$ and $\pi_2(\catid{\Gamma},A)\jdeq
\catid{\ctxext{\Gamma}{A}}$ and let $B\in\mathrm{Ty}(\ctxext{\Gamma}{A})$. Then
we have
\begin{align*}
(\catid{\Gamma})_\ast(A\circ p_B) & \jdeq (\catid{\Gamma})_\ast(A)\circ \pi_2(\catid{\Gamma},A)_\ast(p_B) \\
& \jdeq A\circ p_B
\end{align*}
Also, we have
\begin{align*}
\pi_2(\catid{\Gamma},A\circ p_B) & \jdeq \pi_2(\pi_2(\catid{\Gamma},A),p_B) \\
& \jdeq \pi_2(\catid{\ctxext{\Gamma}{A}},p_B) \\
& \jdeq \catid{\ctxext{{\Gamma}{A}}{B}}
\end{align*}
\item Let $f:\Delta\to \Gamma$ and $g:X\to \Delta$. We will prove by induction
on $A\in\Lambda(\Gamma)$ that $(f\circ g)_\ast(A)\jdeq g_\ast(f_\ast(A))$ and
that $\pi_2(f\circ g,A)\jdeq \pi_2(f,A)\circ \pi_2(g,f_\ast(A))$. Note that
these properties hold by definition for $A\jdeq\catid{\Gamma}$. For the inductive
step, assume that $A\in\Lambda(\Gamma)$ is such that $(f\circ g)_\ast(A)\jdeq g_\ast(f_\ast(A))$ and
$\pi_2(f\circ g,A)\jdeq \pi_2(f,A)\circ \pi_2(g,f_\ast(A))$, and let
$B\in\mathrm{Ty}(\ctxext{\Gamma}{A})$. Then we have
\begin{align*}
(f\circ g)_\ast(A\circ p_B) & \jdeq \pi_2(f\circ g,A)_\ast(p_B) \\
& \jdeq (\pi_2(f,A)\circ\pi_2(g,f_\ast(A)))_\ast(p_B) \\
& \jdeq \pi_2(g,f_\ast)_\ast(\pi_2(f,A)_\ast(p_B)) \\
& \jdeq g_\ast(f_\ast(A\circ p_B))
\end{align*}
\item This property holds by definition.\qedhere
\end{enumerate}
\end{proof}

\begin{defn}
If either of the equivalent conditions of \autoref{thm:Dsys_coh} hold, we say 
that the D-system is \emph{coherent}. Thus, we have a functor
$U:\mathbf{Dsys}\to\mathbf{E'sys_{\circ}}$ from coherent D-systems to E'-systems
\end{defn}

@@

\begin{thm}
The restriction of the functor $U:\mathbf{Dsys}\to\mathbf{E'sys_{\circ}}$ to the
locally well-founded D-systems is full and faithful.
\end{thm}

\begin{proof}
The functor $U$ is faithful on the locally well-founded D-systems, because for any morphism 
$\mathcal{F}:\mathbb{D}\to\mathbb{D}'$ of locally well-founded D-systems, 
$U(\mathcal{F})$ and $\mathcal{F}$ have the same underlying functor
from $\cat{C}$ to $\cat{C}'$.

It remains to show that $U$ is full on the locally well-founded D-systems. Let
$\mathbb{D}$ and $\mathbb{D}$ be locally well-founded D-systems and let $\mathcal{G}:U(\mathbb{D})
\to U(\mathbb{D}')$. Then $\mathcal{G}$ restricted to $\cat{F}$ lifts factorizations.
We  first have to define a natural transformation
$\mathrm{Ty}\Rightarrow \mathrm{Ty}'\circ\mathcal{G}$. Let
$\Gamma\in\cat{C}$ and $A\in\mathrm{Ty}(\Gamma)$. Then we have
$\mathcal{G}(p_A):\mathcal{G}(\ctxext{\Gamma}{A})\to \mathcal{G}(\Gamma)$.
The category of factorizations of $p_A$ in $\cat{F}$ is a isomorphic to $\mathbf{2}$,
so the category of factorizations of $\mathcal{G}(p_A)$ is also isomorphic to
$\mathbf{2}$. This implies that $\mathcal{G}(p_A)$ is of the form $p_B$ for
some $B\in\mathrm{Ty}'(\mathcal{G}(\Gamma))$. This $B$ is called $\mathcal{G}(A)$.
